miniKanren with fair search strategies

### 2019-05-20 Paper/paper.tex
correcting sec. 2

Under the hood, the \conde{} and the call to \texttt{repeato} are connected by 
\conj{}. The \conde{} goal outputs a search space including four trivially 
different states. Then the next conjunctive goal, \texttt{(repeato x q)}, is 
applied to each of these states, producing four trivially different search 
spaces. In the examples above, the \conj{}s are allocating more computational 
resources to the search space of \texttt{a}. On the contrary, fair \conj{} 
would allocate resources evenly to each search space.

@@

A more interesting situation is when the first conjunct produces infinite many 
answers. Consider the following example, a naive specification of fair \conj{} 
might require search strategies to produce all sorts of singleton lists, but no 
longer ones, which makes the strategies incomplete. A search strategy is 
\emph{complete} if and only if it can find out all the answers within finite 
time, otherwise, it is \emph{incomplete}.

@@

BFS puts states of the same cost in the same bag, and allocate resources 
carefully among search spaces related to different bags such that answers are 
produced in increasing order of cost. The \emph{cost} of a answer is its depth 
in the search tree (i.e. the number of calls to relational definitions required 
to find them) \citep{seres1999algebra}. In the following example, the cost of 
each answer is equal to the length of the inner lists plus the length of the 
outer list.

@@

\begin{defn}[fair \conj{}]
A \conj{} is fair if and only if it allocates computational resource evenly to 
search spaces produced from states in the same bag. A bag is a finite 
collection of state. And search strategies with fair \conj{} should represent 
search spaces with possibly infinite collection of state. 
\end{defn}

@@

In this section, we review the implementation of interleaving depth-first 
search 
(\DFSi). This review is for comparison with other search strategies in this 
paper. Thus we focus on parts that are changed later. TRS2 
\citep{friedman_reasoned_2018} provides a comprehensive description of the 
whole miniKanren implementation.

@@

Fig.~\ref{DFSi} shows part of the implementation of \DFSi{}. We follow a
convention to name variables bound to states with `s', to name variables bound 
to goals with `g', and to name variables bound to search spaces with a suffix 
`-inf'. The first function, \texttt{disj2}, implements binary disjunction. 
\texttt{append-inf} is its helper, which composes two disjunctive search 
spaces. 
The following function, \texttt{conj2}, implements binary conjunction. It 
applies 
the \emph{first} goal to the input state, then applies the second goal to 
states in the resulting search space. The latter process is done with a helper 
function. \texttt{append-map-inf} applies its input goal to states in its input 
search spaces and compose the resulting search spaces. It reuses 
\texttt{append-inf} for search space composition. The following three 
definitions introduce syntactic sugars that miniKanren users are more familiar 
with. The first two definitions say disjunction and disjunction are 
right-associative. The next and last definition says \conde{} relates its 
clauses disjunctively, and goals in the same clause conjunctively.

>>>>

Under the hood, the \conde{} and the call to \texttt{repeato} are connected by 
\conj{}. The \conde{} goal outputs a search space including four trivially 
different states. Then the next conjunctive goal, \texttt{(repeato x q)}, is 
applied to each of these states, producing four trivially different search 
spaces. In the examples above, the \conj{}s are allocating more computational 
resources to the search space of \texttt{a}. On the contrary, fair \conj{} 
would allocate resources evenly to each search space. For example,

@@

A more interesting situation is when the first conjunct produces infinite many 
states. Consider the following example, a naive specification of fair \conj{} 
might require search strategies to produce all sorts of singleton lists, but no 
longer ones, which makes the strategies incomplete. A search strategy is 
\emph{complete} if and only if it can find out all the answers within finite 
time, otherwise, it is \emph{incomplete}.

@@

BFS puts states of the same cost in the same bag, and allocate resources 
carefully among search spaces related to different bags such that answers are 
produced in increasing order of cost. The \emph{cost} of a answer is its depth 
in the search tree (i.e. the number of calls to relational definitions required 
to find them) \citep{seres1999algebra}. In the following example, every answer 
is a list of list of symbols, where inner lists are the same. Here the 
cost of each answer is equal to the length of its inner lists plus the length 
of its outer list.

@@

\begin{defn}[fair \conj{}]
A \conj{} is fair if and only if it allocates computational resource evenly to 
search spaces produced from states in the same bag. A bag is a finite 
collection of state. And search strategies with fair \conj{} should represent 
search spaces with possibly infinite collection of bags. 
\end{defn}

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to this paper. TRS2 
\citep{friedman_reasoned_2018} provides a comprehensive description of the 
whole miniKanren implementation.

\begin{figure}
	\lstinputlisting{Figures/conde.rkt}
	\caption{implementation of \conde}
	\label{conde}
\end{figure}

The definition of \conde{} (Fig.~\ref{conde}) is shared by all search 
strategies. It relates clauses disjunctively, and goals in the same 
clause conjunctively.

@@

Fig.~\ref{DFSi} shows parts that are later compared with other search 
strategies. We follow some conventions to name variables: variables bound to 
states are named \texttt{s}; variables bound to goals are named `g' (possibly 
with subscript); variables bound to search spaces have names ending with 
$^\infty$. The first function, \texttt{disj2}, implements binary disjunction. 
\texttt{append-inf} is its helper, which composes two disjunctive search 
spaces. The following function, \texttt{conj2}, implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting search space. The latter process is done with a 
helper function. \texttt{append-map-inf} applies its input goal to states in 
its input search spaces and compose the resulting search spaces. It reuses 
\texttt{append-inf} for search space composition. The following two definitions 
say disjunctive operator and conjunctive operator are right-associative.
